---
title: "CS 584 Final - Fall 2017"
author: "Jake Whalen"
date: "November 21, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source('_Packages.R')
source('_Data.R')
source('_EDA.R')
source('_ModelResults.R')
```

# Executive Summary

Summary of purpose/results

# Motivation

# Data

## Overview
There are +3,000 labelled records. The data contains 20 Features such as mean frequency, skewness, spectral entropy and average of fundamental frequency measured across acoustic signal. The goal is to predict whether a voice is male or female. A follow up goal is to determine what characteristics help distinguish between genders in the dataset.

The data was downloaded from <https://www.kaggle.com/primaryobjects/voicegender>. 



# Approach

EDA

Machine Learning

Evaluation

# Exploratory Data Analysis

#### Classes

```{r, fig.height=3, fig.width=3}
genderCount
```


#### Visualize

###### Distributions

```{r, fig.height=7, fig.width=7}
allDensity
```

###### Boxplot

```{r, fig.height=7, fig.width=7}
allBoxPlot
```

###### Heatmap

```{r, fig.height=7, fig.width=7}
ggheatmap
```

###### Scatter Plot

```{r, fig.height=7, fig.width=7}
oneVall
```

Also discuss 3D

#### Statistical Testing


# Classification

#### Machine Learning

###### Python

###### Scikit Learn

#### K-Nearest Neighbors

Confusion Matrix
```{r, fig.height=3, fig.width=4}
knn_cm + theme(legend.position = 'none')
```

Cross Validation Results
```{r, fig.height=3, fig.width=5}
knn_score.graph
```


#### Decision Tree

Confusion Matrix
```{r, fig.height=3, fig.width=4}
dt_cm + theme(legend.position = 'none')
```

Feature Importance
```{r, fig.height=3, fig.width=4}
dt_feature.graph
```

#### Support Vector Machine

Confusion Matrix
```{r, fig.height=3, fig.width=4}
svm_cm + theme(legend.position = 'none')
```


```{r, fig.height=3, fig.width=4}

```

#### Logistic Regression

Confusion Matrix
```{r, fig.height=3, fig.width=4}

```


```{r, fig.height=3, fig.width=4}

```

#### K-Nearest Neighbors (PCA)

Confusion Matrix
```{r, fig.height=3, fig.width=4}
knn.pca_cm + theme(legend.position = 'none')
```

Cross Validation Results
```{r, fig.height=3, fig.width=5}
knn.pca_score.graph
```

#### Random Forest

Confusion Matrix
```{r, fig.height=3, fig.width=4}
rf_cm + theme(legend.position = 'none')
```


```{r, fig.height=3, fig.width=4}

```

#### Support Vector Machine (PCA)

Confusion Matrix
```{r, fig.height=3, fig.width=4}
svm.pca_cm + theme(legend.position = 'none')
```


```{r, fig.height=3, fig.width=4}

```


# Conclusions



